---
title: "Lab4"
author: "Varun Tripathi"
date: "2024-09-26"
output: "github_document"
---

```{r}
library(car)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)

```

# Load dataset
```{r}

lab4data <- read.csv("C:/Users/vtrip/OneDrive/Documents/GitHub/Lab4/skewed.csv")


```


# Normality
## Normality Plots
```{r}


ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)   + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq() + theme_classic()

qq+ geom_qq_line()


#functions below are base R functions, they can do the job, but not as customizable as ggplot

hist(lab4data$Performance)

qqnorm(lab4data$Performance, col = "steelblue", lwd = 2)

plot(density(lab4data$Performance, na.rm = TRUE, bw = 90),  lwd=2, main = "")

```

### Normality Plots by Group
```{r}

#a fun function to plot the violin plot by group, but it's not part of ggplot
violinBy(Performance ~ Group, data = lab4data, rain= TRUE, vertical = FALSE)


#Find a way to plot the histograms, density, and qq plots by groups using ggplot

ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + facet_wrap(~Group) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2) + facet_wrap(~Group)   + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq() + facet_wrap(~Group)  + theme_classic()

qq+ geom_qq_line()


#Use ggplot to create a violin plot by groups


```


## Normality Tests
```{r}

describe(lab4data$Performance)

shapiro.test(lab4data$Performance)



```

### Normality Tests by Group
```{r}

#Use the describeBy() function to get skewness and kurtosis by group
?describeBy()

describeBy(Performance ~ Group, data = lab4data) 
           
#Use the group by function to get shapiro test results by group
lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance)$statistic, p_value = shapiro.test(Performance)$p.value)

#Use the filter function to get both





```
# Q1: What is your overall conclusion about this variable's normality? Why?

The datasets are not normal as seen in the graphs above - they are skewed. Also, the S-W normality test is significant, further confirming the normality assumption is not met. 

# Equal Variance between Groups
## Descrptive Variance
```{r}

lab4data_clean<-drop_na(lab4data)

var(lab4data_clean$Performance)

lab4data_clean %>%
  group_by(Group) %>%
  summarize(Variance = var(Performance))


```
#reminder that variance is more than double, so not met
## Equal Variance Test
```{r}


leveneTest(Performance~Group, lab4data)

MANOVA(lab4data, dv = "Performance", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 

lab4dataConG1<-lab4data %>%
  filter(Group == "Control" | Group == "G1")

leveneTest(Performance~Group, lab4dataConG1)

#below, you see that its significant, so condition is NOT met


```

# Q2: Overall, does it meet the equal variance assumption across the groups? Why?
No it does not, because the levene test result was significant. ie. the P value was smaller than 0.05

# Transformation
```{r}

#if any of the assumption is not met, use transformation 

lab4data$Performance_log <- log10(lab4data$Performance)

ggplot(lab4data, aes(x = Performance_log)) + geom_histogram(binwidth = 0.2) + theme_classic()

```

# Q3: Run the above tests again with the transformed outcome. Compare the differences in results. 

The histogram for the dataset was originally positively skewed but, after the log transformation, the histogram shows a more normal distribution.