---
title: "Lab4"
author: "Varun Tripathi"
date: "2024-09-26"
output: "github_document"
---

```{r}
library(car)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)

```

# Load dataset
```{r}

lab4data <- read.csv("C:/Users/vtrip/OneDrive/Documents/GitHub/Lab4/skewed.csv")


```


# Normality
## Normality Plots
```{r}


ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)   + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq() + theme_classic()

qq+ geom_qq_line()


#functions below are base R functions, they can do the job, but not as customizable as ggplot

hist(lab4data$Performance)

qqnorm(lab4data$Performance, col = "steelblue", lwd = 2)

plot(density(lab4data$Performance, na.rm = TRUE, bw = 90),  lwd=2, main = "")

```

### Normality Plots by Group
```{r}

#a fun function to plot the violin plot by group, but it's not part of ggplot
violinBy(Performance ~ Group, data = lab4data, rain= TRUE, vertical = FALSE)


#Find a way to plot the histograms, density, and qq plots by groups using ggplot

ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + facet_wrap(~Group) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2) + facet_wrap(~Group)   + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq() + facet_wrap(~Group)  + theme_classic()

qq+ geom_qq_line()


#Use ggplot to create a violin plot by groups


```


## Normality Tests
```{r}

describe(lab4data$Performance)

shapiro.test(lab4data$Performance)



```

### Normality Tests by Group
```{r}

#Use the describeBy() function to get skewness and kurtosis by group
?describeBy()

describeBy(Performance ~ Group, data = lab4data) 
           
#Use the group by function to get shapiro test results by group
lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance)$statistic, p_value = shapiro.test(Performance)$p.value)

#Use the filter function to get both





```
# Q1: What is your overall conclusion about this variable's normality? Why?

The datasets don't seem normal based on the graphs above - they are skewed. Although the normality test by group suggests the normality condition is met, the S-W test for the dataset as a whole had a significant P-value. So, it is justifiable to transform the data based on the skewed plot and the significant P-value for the S-W for the dataset before grouping

# Equal Variance between Groups
## Descrptive Variance
```{r}

lab4data_clean<-drop_na(lab4data)

var(lab4data_clean$Performance)

lab4data_clean %>%
  group_by(Group) %>%
  summarize(Variance = var(Performance))


```
#reminder that variance is more than double, so not met
## Equal Variance Test
```{r}


leveneTest(Performance~Group, lab4data)

MANOVA(lab4data, dv = "Performance", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 

lab4dataConG1<-lab4data %>%
  filter(Group == "Control" | Group == "G1")

leveneTest(Performance~Group, lab4dataConG1)

#below, you see that its significant, so condition is NOT met


```

# Q2: Overall, does it meet the equal variance assumption across the groups? Why?
No it does not, because the levene test result was significant. ie. the P value was smaller than 0.05. In addition, the variances between the groups are more than double in size. 

# Transformation
```{r}

#if any of the assumption is not met, use transformation 

lab4data$Performance_log <- log10(lab4data$Performance)

ggplot(lab4data, aes(x = Performance_log)) + geom_histogram(binwidth = 0.2) + theme_classic()

violinBy(Performance_log ~ Group, data = lab4data, rain= TRUE, vertical = FALSE)

shapiro.test(lab4data$Performance_log)

?describeBy()

describeBy(Performance_log ~ Group, data = lab4data) 

lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance_log)$statistic, p_value = shapiro.test(Performance_log)$p.value)

leveneTest(Performance_log~Group, lab4data)


```

# Q3: Run the above tests again with the transformed outcome. Compare the differences in results. 

The histogram for the dataset was originally positively skewed but, after the log transformation, the histogram shows a more normal distribution. In addition, the S-W test is insignificant, which means the normality condition is not violated - originally, this was not the case. Similarly, the Levine test of variance produced a significant result before the transformation. It is now insignificant, meaning the variance assumption is met. The group level normality test produced the same result as before the transformation - not significant; meaning the normality requirement was met, and is still met.